### 2008-12-19 talk/ecoop2009/clibackend.tex
fig.6 on method and block ids should be changed
changes in reaction of Antonio's comments

\item Each either primary or secondary method implements a certain
  number of blocks, all belonging to the same flow graph. Among these blocks
  there always exists an initial block whose input variables are
  parameters of the method; the input variables of all other blocks
  are local variables of the method. \anto{This is wrong: the signature of the secondary methods is fixed, and input args are passed inside the InputArgs class, not as methodo parameters}
\end{itemize}

@@

Following an internal link is not difficult in IL bytecode: a jump to
the corresponding code fragment in the same method is emitted 
to execute the new block, whereas the appropriate local variables are
used for passing arguments. \anto{this is wrong for the same reason as above}
Also following an external link whose target is the initial block of a
method is not difficult: the corresponding method has to be invoked
with the appropriate arguments.

@@

To solve this problem every method contains a special code, called
\emph{dispatcher}; whenever a method is invoked, its dispatcher is
executed first\footnote{The dispatcher should not be
confused with the initial block of a method.} to
determine which block has to be executed.
This is done by passing to the method a 32 bits number, called 
\emph{block id}, which uniquely identifies the next block of the graph to be executed.
The high word \anto{a word is 32 bit, block num and method id are 16 bit each} of a block id is the id of the method to which the block
belongs, whereas the low word is a progressive number univocally identifying
each block implemented by the method.

The picture in Figure~\ref{block-id-fig} shows a graph composed of three methods (for
simplicity, dispatchers are not shown); method ids are in red, whereas
block numbers are in black. 
The graph contains three external links; in particular, note the link
between blocks \texttt{0x00020001} and \texttt{0x00010001} which
connects two blocks implemented by different methods.
\begin{figure}[h]
\begin{center}
\includegraphics[height=6cm]{blockid}
\caption{Method and block ids.}\label{block-id-fig}
\end{center}
\end{figure}

For instance, the code\footnote{For simplicity we write C\# code instead of
the actual IL bytecode.} generated for the dispatcher of method \texttt{0x0002}
is similar to the following fragment: 
\begin{small}
\begin{lstlisting}[language={[Sharp]C}]
// dispatch block
int methodid = (blockid && 0xFFFF0000) >> 16; 
int blocknum = blockid && 0x0000FFFF;         
if (methodid != MY_METHOD_ID) {
  // jump_to_ext 
  ...
}
switch(blocknum) {
  case 0: goto block0;
  case 1: goto block1;
  default: throw new Exception("Invalid block id");
}
\end{lstlisting}
\end{small}
If the next block to be executed is implemented in the same method
({\small\lstinline{methodid == MY_METHOD_ID}}), then the appropriate
jump to the corresponding code is executed, hence internal links
can be managed efficiently. \anto{wrong: internal links don't go through the dispatcher}
Otherwise, the \lstinline{jump_to_ext}
part of the dispatcher has to be executed.
The code that actually jumps to an external block is contained in
the dispatcher of the primary method, whereas the
\lstinline{jump_to_ext} code of dispatchers of secondary methods
simply delegates the dispatcher of the primary method of the same
graph (see later).

>>>>

\item Each either primary or secondary method implements a certain
  number of blocks, all belonging to the same flow graph. Among these blocks
  there always exists an initial block whose input arguments 
  might be passed as arguments of the method; however, for
  implementation  reasons (see the details below) the input variables
  of all blocks (including the initial one)
  are implemented as local variables of the method. 
\end{itemize}

@@

Following an internal link would  not be difficult in IL bytecode: a jump to
the corresponding code fragment in the same method can be emitted 
to execute the new block, whereas the appropriate local variables can be
used for passing arguments. 
Also following an external link whose target is an initial block could
be easily implemented, by just invoking the corresponding method.

@@

To solve this problem every method contains a special code, called
\emph{dispatcher}; whenever a method is invoked, its dispatcher is
executed first\footnote{The dispatcher should not be
confused with the initial block of a method.} to
determine which block has to be executed.
This is done by passing to the method a 32 bits number, called 
\emph{block id}, which uniquely identifies the next block of the graph to be executed.
The high 2 bytes \dacom{word was meant as a fixed-sized group of bits} of a block id constitute the id of the method to which the block
belongs, whereas the low 2 bytes constitute a progressive number univocally identifying
each block implemented by the method.

The picture in Figure~\ref{block-id-fig} shows a graph composed of three methods (for
simplicity, dispatchers are not shown); method ids are in red, whereas
block numbers are in black. 
The graph contains three external links; in particular, note the link
between blocks \texttt{0x00020001} and \texttt{0x00010001} which
connects two blocks implemented by different methods.
\begin{figure}[h]
\begin{center}
\includegraphics[height=6cm]{blockid}
\caption{Method and block ids.\dacom{in the picture ``Main method'' must be
    replaced with ``Primary method'' and in the primary method the block with number 0 should be added}}\label{block-id-fig}
\end{center}
\end{figure}

For instance, the code\footnote{For simplicity we write C\# code instead of
the actual IL bytecode.} generated for the dispatcher of method \texttt{0x0002}
is similar to the following fragment: 
\begin{small}
\begin{lstlisting}[language={[Sharp]C}]
// dispatch block
int methodid = (blockid && 0xFFFF0000) >> 16; 
int blocknum = blockid && 0x0000FFFF;         
if (methodid != MY_METHOD_ID) {
  // jump_to_ext 
  ...
}
switch(blocknum) {
  case 0: goto block0;
  case 1: goto block1;
  default: throw new Exception("Invalid block id");
}
\end{lstlisting}
\end{small}
If the next block to be executed is implemented in the same method
({\small\lstinline{methodid == MY_METHOD_ID}}), then the appropriate
jump to the corresponding code is executed.
Otherwise, the \lstinline{jump_to_ext}
part of the dispatcher has to be executed.
The code that actually jumps to an external block is contained in
the dispatcher of the primary method, whereas the
\lstinline{jump_to_ext} code of dispatchers of secondary methods
simply delegates the dispatcher of the primary method of the same
graph (see later).
